---
title: 'Assignment #6'
author: "Yutae Lee (626005947)"
date: "2022-10-17"
output:
  pdf_document:
    latex_engine: xelatex
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 6.1 

**a)**


$$cov(\theta_A,\theta_B) = E[\theta_A \theta_B] * E[\theta_A]E[\theta_B]$$
Given that $\theta_A = \theta, \theta_B = \theta * \gamma$

We can see that:

$$cov(\theta_A, \theta_B) = E[\theta^2 * \gamma] -E[\theta] E[\theta * \gamma]$$

$$=E[\theta^2]E[\gamma] - E[\theta]^2E[\gamma]$$

$$=(E[\theta^2] - E[\theta]^2)E[\gamma]$$

$$= var(\theta)E[\gamma] \neq 0$$


This indicates that $\theta_A$, $theta_B$ are dependent.

When $\theta_B$ is a product of $\theta_A$ + Gamma distributed noise, then this joint prior would justified.

**b)**

The joint prior distribution is:

$$p(\theta,\gamma|\textbf{y}_A,\textbf{y}_B)\propto p(\theta,\gamma) * p(\textbf{y}_A,\textbf{y}_B|\theta,\gamma)$$
$$= p(\theta)p(\gamma) * p(\textbf{y}_A|\theta) * p(\textbf{y}_B|\theta) $$

$$\propto \left(\theta^{a_\theta - 1}e^{-b_\theta \theta}\right) * \left(\gamma^{a_\gamma - 1}e^{-b_\gamma \gamma} \right) *  \left(\prod_{i=1}^{n_{A}} \theta^{y_{A_i}} e^{-\theta} \right) * \left(\prod_{i=1}^{n_{B}} (\gamma \theta)^{y_{B_i}} e^{-\gamma \theta} \right) \\$$
$$= \left(\theta^{a_\theta - 1}e^{-b_\theta \theta}\right) * \left(\gamma^{a_\gamma - 1}e^{-b_\gamma \gamma} \right) *  \left( \theta^{\sum_{i = 1}^{n_A} y_{A_i}} e^{-n_A \theta} \right) * \left( (\gamma \theta)^{\sum_{i=1}^{n_B} y_{B_i}} e^{- n_B \gamma \theta} \right) \\$$
$$= \left(\theta^{a_\theta - 1}e^{-b_\theta \theta}\right) * \left(\gamma^{a_\gamma - 1}e^{-b_\gamma \gamma} \right) *  \left( \theta^{n_A \bar{y}_A} e^{-n_A \theta} \right) * \left( (\gamma \theta)^{n_B \bar{y}_B} e^{- n_B \gamma \theta} \right)$$


From this we can say that:

$$p(\theta, \mid \textbf{y}_A, \textbf{y}_B, \gamma) \propto \left(\theta^{a_\theta - 1}e^{-b_\theta \theta}\right) * \left(\gamma^{a_\gamma - 1}e^{-b_\gamma \gamma} \right) *  \left( \theta^{n_A \bar{y}_A} e^{-n_A \theta} \right) * \left( (\gamma \theta)^{n_B \bar{y}_B} e^{- n_B \gamma \theta} \right)$$
$$\propto \left(\theta^{a_\theta - 1}e^{-b_\theta \theta}\right) * \left( \theta^{n_A \bar{y}_A} e^{-n_A \theta} \right) * \left( (\gamma \theta)^{n_B \bar{y}_B} e^{- n_B \gamma \theta} \right) $$

$$\propto \theta^{a_\theta + n_A \bar{y}_A + n_B \bar{y}_B - 1} \exp \left( - (b_\theta + n_A + n_B \gamma ) \theta \right)$$


$$\propto dgamma\left(a_\theta + n_A \bar{y}_A + n_B \bar{y}_B, b_\theta + n_A + n_B \gamma \right)$$ 

**c)**

$$p(\gamma, | \textbf{y}_A, \textbf{y}_B, \theta)$$
$$\propto \left(\theta^{a_\theta - 1}e^{-b_\theta \theta}\right) * \left(\gamma^{a_\gamma - 1}e^{-b_\gamma \gamma} \right) *  \left( \theta^{n_A \bar{y}_A} e^{-n_A \theta} \right) * \left( (\gamma \theta)^{n_B \bar{y}_B} e^{- n_B \gamma \theta} \right) \\$$
$$\propto \left(\gamma^{a_\gamma - 1}e^{-b_\gamma \gamma} \right) * \left( (\gamma \theta)^{n_B \bar{y}_B} e^{- n_B \gamma \theta} \right)$$
$$\propto \left(\gamma^{a_\gamma - 1}e^{-b_\gamma \gamma} \right) * \left( \gamma^{n_B \bar{y}_B} e^{- n_B \gamma \theta} \right) \\$$
$$\propto \gamma^{a_\gamma + n_B \bar{y}_B - 1} \exp\left( -(b_\gamma + n_B \theta) \gamma \right)$$

$$\propto \text{dgamma}\left(a_\gamma + n_B\bar{y}_B, b_\gamma + n_B \theta \right)$$


**d)**


```{r}
library(ggplot2)
Y_a <- scan(url('https://www2.stat.duke.edu/courses/Fall09/sta290/datasets/Hoffdata/menchild30bach.dat'))
Y_b <- scan(url('https://www2.stat.duke.edu/courses/Fall09/sta290/datasets/Hoffdata/menchild30nobach.dat'))
n_a = length(Y_a)
n_b = length(Y_b)
ybar_a = mean(Y_a)
ybar_b = mean(Y_b)
a_theta = 2
b_theta = 1
S = 5000
ab_gamma = c(8, 16, 32, 64, 128)
theta_diff = sapply(ab_gamma, function(abg) {
  a_gamma = b_gamma = abg
  THETA = numeric(S)
  GAMMA = numeric(S)
  theta = ybar_a
  gamma = ybar_a / ybar_b  
  for (s in 1:S) {
    theta = rgamma(
      1,
      a_theta + n_a * ybar_a + n_b * ybar_b,
      b_theta + n_a + n_b * gamma
    )
    
    gamma = rgamma(
      1,
      a_gamma + n_b * ybar_b,
      b_gamma + n_b * theta
    )
    THETA[s] = theta
    GAMMA[s] = gamma
  }

  THETA_A = THETA
  THETA_B = THETA * GAMMA
  mean(THETA_B - THETA_A)
})
ggplot(data.frame(ab_gamma = ab_gamma, theta_diff = theta_diff), aes(x = ab_gamma, y = theta_diff)) +
  geom_point() +
  geom_line()
```



The gamma distribution is centered around 1 and the magnitude is how strong is our belief in which $\theta_B / \theta_A$ is 1. We can see that as the number of $\theta_B / \theta_A$ increases, the difference decreases.







