---
title: 'Assignment #2'
author: "Yutae Lee (626005947)"
date: "2022-09-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 3.1

**a)**

$$P(Y_1,......,Y_100|\theta) = \theta^{\sum_{i = 1}^{100}{y_i}}({1-\theta})^{100 - \sum _ {i=1}^{100}{y_i}}$$
$$Pr(\sum{Y_i} = y|\theta) = {100 \choose y}\theta^y(1-\theta)^{100-y}$$

**b)**

$$Pr(\sum{Y_i} = 57| \theta) =  {100 \choose 57}\theta^{57}(1-\theta)^{43}$$

```{r}
theta=seq(0,1,by = 0.1)
probability = (choose(100,57)) * (theta ^ 57) * ((1-theta) ^ 43)
probability
plot(theta,probability,type = 'h', ylab = 'Probability')
```


**c)**

$$P(\theta|y) =  \frac{{100 \choose 57} \theta^{57} (1-\theta)^{43}(1/11)}{p(y)} \propto \theta^{57}(1-\theta)^{43}(1/11) = c * \theta^{57}(1-\theta)^{43}(1/11) $$

```{r}
prior_probability = rep(1/11,11)
post_probability = theta ^ 57 * (1-theta) ^ 43 * (1/11)
post_probability = post_probability/sum(post_probability)
post_probability
plot(theta,post_probability, type = 'h', ylab = "Posterior Probability")

```




**d)**


```{r}
theta = seq(0,1,length = 200)
prior_probability = 1
posterior_probability = choose(100,57) * theta ^ 57 * (1-theta) ^ 43 * 1
posterior_probability
plot(theta,posterior_probability,type = 'l' , ylab = 'Posterior Probability')
```



**e)**


```{r}
theta = seq(0,1,length = 200)
posterior_probability = dbeta(theta,58,44)
posterior_probability
plot(theta,posterior_probability,type = 'l', ylabs = "Posterior Probability")
```

```{r}
theta[posterior_probability == max(posterior_probability)]
```
Using the beta distribution we are able to get more accuracy in estimating $\theta$. We can also see that all of the graph shows similar values  within themselves. Unlike other priors which gave us $\theta = 0.6$, in beta we can get $\theta = 0.5678392$.



## Problem 3.3

**a)**

$x_A$ ~ $poi(\theta_A)$, $x_B$ ~ $poi(\theta_B)$

$\theta_A$ ~ $\gamma(120,10)$, $\theta_B$ ~ $\gamma(12,1)$

$f(\theta_A) = \frac{10^{120}}{\sqrt{120}} e ^{-10\theta_A} * \theta_A ^ {120-1}$
$\beta(x_A) = e^{-\theta_A}\frac{\theta_A^{x_A}}{x_A!}$

$Y_A = \sum_{i = 1}^{10} {{x_A}_i}$~$poi(10\theta_A)$
$\beta(y_A) = e^{-\theta_A}\frac{(10\theta_A)^{y_A}}{y_A!}$

$f(y_A,\theta_A) = \frac{10^{120}}{\sqrt{120}}e^{-10\theta_A}\theta_A^{119}*\frac{e^{-10\theta_A}*(10\theta_A)^{y_A}}{y_A!}$

$f(y_A) = \int_{0}^{\inf}f(y_A,\theta_A)d\theta_A$

$=\frac{10^{120}}{\sqrt{120}y_A!}* 10^{y_A}\int_{0}^{inf}e^{-20}\theta_A\theta_A^{120+y_A-1}d\theta_A$

$= \frac{10^{120}10y_A}{\sqrt{120}y_A!}* \frac{\sqrt{120+y_A}}{20^{120+y_A}}$

$f(\theta_A|y_A) = \frac{f(y_A*\theta_A)}{f(y_A)} = \frac{e^{-20}\theta_A\theta_A^{120y_A-1}}{\sqrt{120+y_A}/(20^{120+y_A})}$

~ $\gamma(120+y_A,20)$

= $\gamma(120+117,20) = \gamma(237,20)$

mean = 11.85

var = 0.5929

95% C.I. = (10.38925,13.40545)

Doing the same process for $\theta_B$ I get $\gamma(125,19)$

I get mean = 8.9286

Var = 0.0.6378

95% C.I. = (7.432064,10.56031)


Sorry for the unfinished work but this is all I have upto.